Concurrent Programming with CXL

### 2020-06-23 book/paper.tex
Latest

\section*{Exercises}
\begin{problems}
\item Implement a version of parallel qsort (see Example~\ref{ex:qsort})
using the actor model.  Use an actor for each worker as well as a main
actor that waits for each worker actor to have finished.
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\end{problems}

@@

Method \texttt{produce(item)} uses \texttt{inc \&(back)} to allocate
the next available slot in the \texttt{items} array.
It stores the item as a singleton tuple.
Method \texttt{consume()} repeatedly scans the array, up to the
\texttt{back} index, trying to find an item to return.
To check an entry, it uses \texttt{swap()}
to atomically remove an item from a slot if there is one.
This way, if two or more processes attempt to extract an item from
the same slot, at most one will succeed.

@@

The solution is not practical because the \texttt{items} array must be
unbounded if an unbounded number of items may be produced.  Each slot
in the array is only used once, which is inefficient.  Also, the \texttt{inc}
and \texttt{swap} interlock instructions are not universally available
on existing hardware.
However, in the literature there are many examples of practical
non-blocking (aka \emph{wait-free})
\index{wait-free synchronization}
synchronization algorithms~\cite{}.

>>>>

\section*{Exercises}
\begin{problems}
\item Actors and message queues are good for building pipelines.
Develop a pipeline that computes Mersenne primes (primes that are one less
than a factor of two).  Write four actors:
\begin{enumerate}
\item an actor that generates a sequence of integers 1 through \texttt{N};
\item an actor that receives integers and forwards only those that are prime;
\item an actor that receives integers and forwards only those that are one
less than a factor of two;
\item an actor that receives integers but otherwise ignores them.
\end{enumerate}
Configure two versions of the pipeline, one that first checks if a number
is prime and then if it is one less than a factor of two, the other
in the opposite order.  Which do you think is better?
\item Bonus question: implement a version of parallel qsort
(see Example~\ref{ex:qsort}) using the actor model.
\end{problems}

@@

Method \texttt{produce(item)} uses \texttt{inc(\&back)} to allocate
the next available slot in the \texttt{items} array.
It stores the item as a singleton tuple.
Method \texttt{consume()} repeatedly scans the array, up to the
\texttt{back} index, trying to find an item to return.
To check an entry, it uses \texttt{swap()}
to atomically remove an item from a slot if there is one.
This way, if two or more processes attempt to extract an item from
the same slot, at most one will succeed.

@@

The solution is not practical because the \texttt{items} array must be
unbounded if an unbounded number of items may be produced.  Each slot
in the array is only used once, which is inefficient.  Also, the \texttt{inc}
and \texttt{swap} interlock instructions are not universally available
on existing hardware.
However, in the literature there are many examples of practical
non-blocking (aka \emph{wait-free})
\index{wait-free synchronization}
synchronization algorithms~\cite{}.
There are also various algorithms where read-only access does not
require a lock but updates do require a lock---these are useful in
situations where most accesses are read-only.
